Collective behavior of Janus Particles Suspended in a Viscous Fluid

### 2022-08-05 doc/PRF2022/PRF_main.tex
added refs to intro; updated references.bib

\begin{abstract}
%<<<<<<< HEAD
%Active colloidal systems with non-equilibrium self-organization is a long-standing challenge in biology. To understand how hydrodynamic flow may be used
%to actively control self-assembly of Janus particles (JPs),  we use a model recently developed for the many-body hydrodynamics of amphiphilic JPs suspended 
%in a viscous background flow (JFM, 941, 2022) to investigate how various bilayer structures arise from tuning the hydrophobicity/hydrophilicity of the JPs. 
%Focusing on three distributions of hydrophobicity we found JPs may assemble into uni-lamella, multi-lamella, and striated structures in a viscous fluid. 
%Under a linear flow and a Taylor-Green mixing flow,  we use three measures to quantify the collective dynamics of JP particles under a background flow: 
%(a) Free energy from the hydrophobic interactions between the JPs, (b) an order parameter for the ordering of JPs in terms of alignment of their directors, 
%and (c) a strain parameter that captures the deformation in the assembly. We found the dynamics of these three measures correlate well with the hydrodynamics 
%of JPs. These numerical provide insights into dynamic control of non-equilibrium active biological systems with similar self-organization.
% insert abstract here
%=======
  Active colloidal systems with non-equilibrium self-organization is a
  long-standing, challenging area in biology. To understand how
  hydrodynamic flow may be used to actively control self-assembly of
  Janus particles (JPs), we use a model recently developed for the
  many-body hydrodynamics of amphiphilic JPs suspended in a viscous
  background flow (JFM, 941, 2022). We investigate how various bilayer
  structures arise from tuning the hydrophobic properties of the
  JP-solvent interface. Focusing on a simple tuning of the
  hydrophobicity distribution, we found JPs assembled into uni-lamella,
  multi-lamella and striated structures. To introduce dynamics, we
  include a linear shear flow and a Taylor-Green mixing flow, and
  measure the collective dynamics of JP particles in terms of their (a)
  free energy from the hydrophobic interactions between the JPs, (b)
  order parameter for the ordering of JPs in terms of alignment of their
  directors, and (c) strain parameter that captures the deformation in
  the assembly. We characterized the effective material properties of
  the JP structures and found that the uni-lamellar structures increases
  orientation order under shear flow, the multilamellar structure
  behaves as a shear thinning fluid, and the striated structure
  possesses a finite yield stress. These numerical result provide
  insights into dynamic control of non-equilibrium active biological
  systems with similar self-organization.
%>>>>>>> 4f49b2ee420baba836cbb337c263b33981e1b096
\end{abstract}

@@

\section{Introduction}
% Put \label in argument of \section for cross-referencing
%\section{\label{}}
Janus particles are colloids whose surfaces 
%have different properties. For example, it is common for the surface of a Janus
%particle to be 
are partly hydrophilic and hydrophobic (polar and non-polar).
More generally the surface of a Janus particle may have a spatial distribution of hydrophobicity/hydrophilicity.
In the absence of any imposed flow, Janus particles often self-assemble into
structures in a viscous solvent with tunable functions for biomedical engineering
applications~\cite{GheisariSahfieeAbbasiEtAl2021_DMR,
LiuYangHuangEtAl2016_Angew, LiWangYaoEtAl2019_Nanoscale, Bradley2016}.
Depending on the interactions between Janus particles (JPs) and the
surrounding viscous solvent, JPs self-assemble into oligomers of various
geometries and sizes~\cite{Bradley2017}.

When driven by an external flow, 
dynamic rearrangement of JPs emerges naturally from the interactions between particles and fluid in colloidal
matter~\cite{Brandner2019,RevModPhys.93.025008}. Such colloidal hydrodynamics belongs to a wide class of
nonequilibrium self-organization in physics, often with complexity and features similar to that of 
biological systems such as living cells, bacterial baths, and
animal flocks~\cite{CollardGrosjeanVandewalle2020,Vutukuri2020}. A
long-standing challenge in fluid mechanics and material science is to
solve the ``inverse" problem of creating a model colloidal system that
will self-assemble into a prescribed
structure~\cite{PhysRevLett.128.256102}. For example, the framework of
geometrical frustration, which has been used to explain disordered
systems, could instead be used to design new, ordered systems, through
specific choices for the shapes or interactions of the
particles~\cite{Manoharan2015_Science}.

The many-body hydrodynamics of amphiphilic Janus particles assembled as
vesicles suspended in a viscous fluid in the inertialess regime (zero Reynolds number) has
been studied using boundary integral numerical
simulations~\cite{Fu20,Fu2022_JFM}. The dynamics of the JP suspension
results from the combination of long-range hydrodynamic interaction and
non-local interactions between JPs through the distribution of a
hydrophobic attraction potential (HAP).
%
In a quiescent flow, numerical simulations of a JP suspension showed
self-assembly into micelles and bilayers of JPs that provide an
alternative means for computing the mechanical moduli of a colloidal
membrane~\cite{NaTr00,Fu20,KrFiGuKaHa13}. Under background flows, the
hydrodynamics of a JP vesicle (a self-enclosed bilayer of JPs) exhibit
many familiar behaviors of a vesicle: elongation and
alignment along the extensional direction, tank-treading, and rupture of
a vesicle under shear flow~\cite{Fu2022_JFM}.

In this work, we take advantage of the flexibility of the HAP model to
examine the effects of varying hydrophobicity/hydrophilicity on JP surfaces.
Such variation of the boundary condition on JP surfaces may be realized in the experiments by using
chemicals to adjust the polarity of the viscous solvent, for example ({\bf references needed here}).
We show that a simple tuning of the hydrophobic distribution leads to transitions from unilamellar to
multilamellar or striated bilayer assembly (structures) of JPs. 
Focusing on the fluid-structure interactions that correspond to such transitions, 
we investigate the deformation of these
novel structures in background flows and map out their collective
behavior away from equilibrium.

@@

We set $\max g>0$ to guarantee at least partial hydrophobicity of the JP
surface.  Then, the boundary condition (BC) can be classified into three
categories: $\min g(\xx) = 0$, $\min g(\xx) >0$, and $\min g(\xx) <0$
where the extrema are taken for $\xx \in \Gamma_i$.  These three
categories correspond to three characteristic values of $b$ in
\eqref{eq:bc-type}: $b=1$, $b=2$ and $b=0$, respectively. 
%
To see the effect of the shift parameter $b$, we simulate the dynamics
of 198 particles and 60 particles in a quiescent flow as shown in
Figure~\ref{fig:relax}. Three distinct configurations emerge: (i)
bilayer for $b = 1$, (ii) multilamellar for $b = 2$, and (iii) striated
for $b = 0$.

@@

In BC (ii) with $b = 2$, the boundary data is everywhere positive. It
takes the maximum value $3a$ on the tail (red) and the minimum value $a$
on the head side (yellow). Both sides of the Janus particle are
hydrophobic but more so on the $\alpha = 0$ side. Over a short time
scale, these particles also self-assemble into bilayers. But unlike for
BC (i), the heads are also hydrophobic and so in the long-time dynamics
these bilayers form multilamella structures~\cite{C9NR05885K}. The
number of folds depends primarily on the number of particles.
Figure~\ref{fig:relax}(c), for example, shows a multilamellar structure
with four layers and one with two layers in the inset.

@@

Finally, we use the scalar order parameter for orientational order:
\begin{equation}
  \label{eq:S2}
S_2 = \frac{1}{N_b} \sum_{i=1}^{N_b} \frac{1}{2}(3\cos^2(\theta_i - \bar \theta) - 1).
\end{equation}
Here $\bar \theta$ is a circular mean defined as the orientation of the
principal eigenvector of the matrix $\dd_1\dd_1^\top + \cdots +
\dd_{N_b}\dd_{N_b}^\top$. A value $S_2(t) = 1$ indicates that all
particle directors lie on a common axis e.g., two parallel vectors with
possibly opposite direction are ordered. A value far from 1 indicates
disorder.

@@

We can characterize the effective material properties of the JP
structures based on the changes in free energy, orientational order, and
strain as a function of flow strength. Generally, $F$ increases, $\tilde
S_2$ decreases, and $E$ increases with increases in both $\dot \gamma$
and $V_0$, as expected. Structurally, the striated JP configuration for
BC (iii) behaved the most like a rigid body over the flow strengths
tested. For this structure, the deformation measures were constant in
time for shear rates $\dot \gamma$ up to $0.15$ and TG flow rate $V_0$
up to $0.25$ (Figure~\ref{fig:BC3_shear} and Figure~\ref{fig:BC3_TG}).
Conversely, a $25\%$--$50 \%$ smaller value for $\dot \gamma$ or $V_0$ was
required to observe nonrigid deformations for the BC (i) and BC (ii)
cases.

@@

The characteristics of the self-assembly of amphiphilic JP into
onion-like dendrimersomes like for BC (ii) were previously studied by molecular dynamics
simulations~\cite{C9NR05885K}. The molecular dynamics simulations use an
anisotropic pair potential to describe the particle interactions
where a harmonic form describes the repulsive part and an anisotropic
attractive part describes the hydrophobic interactions.

@@

The results reported here provide an inroad modeling framework for hydrodynamics of active colloids. 
The present study also helps us understand the rheology of JP oligomers that may be realized in the experiments.
We are now extending this study to three dimensional systems with more realistic features such as size distributions of JPs
and thermal fluctuations.
%
 %it has been included in our future goals\cite{kohl-cor-che-vee22}.
From a numerical perspective,
it is straightforward to include
random perturbations in the particle shape and
boundary condition that mimick interfacial properties
found in lab conditions\cite{Bradley2016,Bradley2017}.

>>>>

\begin{abstract}
  Active colloidal systems with non-equilibrium self-organization is a
  long-standing, challenging area in biology. To understand how
  hydrodynamic flow may be used to actively control self-assembly of
  Janus particles (JPs), we use a model recently developed for the
  many-body hydrodynamics of amphiphilic JPs suspended in a viscous
  background flow (JFM, 941, 2022). We investigate how various 
  morphologies arise from tuning the hydrophobic distribution of the
  JP-solvent interface.  
  We found JPs assembled into uni-lamella,
  multi-lamella and striated structures. To introduce dynamics, we
  include a linear shear flow and a Taylor-Green mixing flow, and
  measure the collective dynamics of JP particles in terms of their (a)
  free energy from the hydrophobic interactions between the JPs, (b)
  order parameter for the ordering of JPs in terms of alignment of their
  directors, and (c) strain parameter that captures the deformation in
  the assembly. We characterized the effective material properties of
  the JP structures and found that the uni-lamellar structures increases
  orientation order under shear flow, the multilamellar structure
  behaves as a shear thinning fluid, and the striated structure
  possesses a yield stress. These numerical result provide
  insights into dynamic control of non-equilibrium active biological
  systems with similar self-organization.
\end{abstract}

@@

\section{Introduction}
% Put \label in argument of \section for cross-referencing
%\section{\label{}}
Janus particles are colloids combining two dissimilar chemical or
physical functionalies at their opposite sides\cite{KirillovaMarschelkeSynytska2019}.
Self-propelling Janus particles, for example, with a permanent biphasic asymmetry,
have emerged as a rich chemical platform for the exploration of active matter\cite{Meredithetal2022}.
In the absence of mobility and any imposed flow, Janus particles often self-assemble into
structures in a viscous solvent with tunable functions for biomedical engineering
applications~\cite{GheisariSahfieeAbbasiEtAl2021_DMR,
LiuYangHuangEtAl2016_Angew, LiWangYaoEtAl2019_Nanoscale, Bradley2016,Zarzaretal2015,KirillovaMarschelkeSynytska2019}.
Depending on the interactions between Janus particles (JPs) and the
surrounding viscous solvent, JPs self-assemble into oligomers of various
geometries and sizes~\cite{Bradley2017,KangHonciuc2018,HongCacciutoLuijtenGranick2008}.

When driven by an external flow, 
dynamic rearrangement of JPs emerges naturally from the interactions between particles and fluid in colloidal
matter~\cite{RevModPhys.93.025008}.
Such colloidal hydrodynamics belongs to a wide class of
nonequilibrium self-organization in physics, often with complexity and features similar to that of 
biological systems such as living cells, bacterial baths, and
animal flocks~\cite{CollardGrosjeanVandewalle2020,Vutukuri2020}. A
long-standing challenge in fluid mechanics and material science is to
solve the ``inverse" problem of creating a model colloidal system that
will self-assemble into a prescribed
structure~\cite{PhysRevLett.128.256102}. For example, the framework of
geometrical frustration, which has been used to explain disordered
systems, could instead be used to design new, ordered systems, through
specific choices for the shapes or interactions of the
particles~\cite{Manoharan2015_Science}.

The surface of a Janus particle may have
a dissimilar spatial distribution of hydrophobicity\cite{doi:10.1021/la503455h}.
The many-body hydrodynamics of amphiphilic Janus particles assembled as
vesicles suspended in a viscous fluid in the inertialess regime (zero Reynolds number) has
been studied using boundary integral numerical
simulations~\cite{Fu20,Fu2022_JFM}. The dynamics of the JP suspension
results from the combination of long-range hydrodynamic interaction and
non-local interactions between JPs through the distribution of a
hydrophobic attraction potential (HAP).
%
In a quiescent flow, numerical simulations of a JP suspension showed
self-assembly into micelles and bilayers of JPs that provide an
alternative means for computing the mechanical moduli of a colloidal
membrane~\cite{NaTr00,Fu20,KrFiGuKaHa13}. Under background flows, the
hydrodynamics of a JP vesicle (a self-enclosed bilayer of JPs) exhibit
many familiar behaviors of a vesicle: elongation and
alignment along the extensional direction, tank-treading, and rupture of
a vesicle under shear
flow~\cite{Fu2022_JFM,grandmaison_brancherie_salsac_2021,D2SM00179A,keller_skalak_1982,Finken08,Shaqfeh11}.

Molecular dynamics (MD) and Monte Carlo provide another avenue for
simulating the interaction between JPs, solvents, and
substrates\cite{Brandner2019,Baniketal2021,HongCacciutoLuijtenGranick2008,C9NR05885K}.
These methods often use pair-potentials to describe the
interaction between JPs, for example by prescribing angle
dependent forces and torques that bring the
hydrophobic sides of two amphiphilic JPs into opposition.
The HAP formulation of the present work also has attractive, long-range forces and torques,
but they are instead derived from a boundary value problem for the molecular
structure of water\cite{Ma77, GoHaKo94,ErLjCl89,Lietal05,Israelachvili80}.
Unlike in MD or Monte Carlo-based simulations,
the HAP interactions are nonadditive\cite{Fu20}, so that
the interactions between a pair of JP is affected by the presence of other particles.

In this work, we take advantage of the flexibility of the HAP model to
examine the effects of varying the distribution of hydrophobicity on JP surfaces.
Such variation of the boundary condition on JP surfaces have been realized experimentally
by using chemicals to adjust the polarity of the viscous solvent, for example
\cite{Zarzaretal2015,KirillovaMarschelkeSynytska2019,doi:10.1021/la503455h}.
We show that a simple tuning of the hydrophobic distribution leads to transitions from unilamellar to
multilamellar or striated superstructures of JPs. 
Focusing on the fluid-structure interactions that correspond to such transitions, 
we investigate the deformation of these
novel structures in background flows and map out their collective
behavior away from equilibrium.
Looking forward, including other fields like
electric potential for JPs synthesized with
charged polymers\cite{HongCacciutoLuijtenGranick2008,doi:10.1021/la503455h,KangHonciuc2018},
is straightforward within the context of the boundary integral representations\cite{kohl-cor-che-vee22},
opening further lines of investigation.

@@

We set $\max g>0$ to guarantee at least partial hydrophobicity of the JP
surface.  Then, the boundary condition (BC) can be classified into three
categories: $\min g(\xx) = 0$, $\min g(\xx) >0$, and $\min g(\xx) <0$
where the extrema are taken for $\xx \in \Gamma_i$.  These three
categories correspond to three characteristic values of $b$ in
\eqref{eq:bc-type}: $b=1$, $b=2$ and $b=0$, respectively. 
%
To see the effect of the shift parameter $b$, we simulate the dynamics
of 198 particles and 60 particles in a quiescent flow as shown in
Figure~\ref{fig:relax}. Three distinct configurations emerge: (i)
bilayer for $b = 1$, (ii) multilamellar for $b = 2$, and (iii) striated
for $b = 0$.
Supplementary Movie S1 shows the self-assembly process for the three cases.

@@

In BC (ii) with $b = 2$, the boundary data is everywhere positive. It
takes the maximum value $3a$ on the tail (red) and the minimum value $a$
on the head side (yellow). Both sides of the Janus particle are
hydrophobic but more so on the $\alpha = 0$ side. Over a short time
scale, these particles also self-assemble into bilayers. But unlike for
BC (i), the heads are also hydrophobic and so in the long-time dynamics
these bilayers form multilamella structures~\cite{C9NR05885K}. The
number of layers depends primarily on the number of particles.
Figure~\ref{fig:relax}(c), for example, shows a multilamellar structure
with four layers and one with two layers in the inset.

@@

Finally, we use the scalar order parameter for orientational order\cite{Selinger2016}:
\begin{equation}
  \label{eq:S2}
S_2 = \frac{1}{N_b} \sum_{i=1}^{N_b} \frac{1}{2}(3\cos^2(\theta_i - \bar \theta) - 1).
\end{equation}
Here $\bar \theta$ is a circular mean defined as the orientation of the
principal eigenvector of the matrix $\dd_1\dd_1^\top + \cdots +
\dd_{N_b}\dd_{N_b}^\top$. A value $S_2(t) = 1$ indicates that all
particle directors lie on a common axis e.g., two parallel vectors with
possibly opposite direction are ordered. A value far from 1 indicates
disorder.

@@

We can characterize the effective material properties of the JP
structures based on the changes in free energy, orientational order, and
strain as a function of flow strength. Generally, $F$ increases, $\tilde
S_2$ decreases, and $E$ increases with increases in both $\dot \gamma$
and $V_0$, as expected. Structurally, the striated JP configuration for
BC (iii) behaved the most like a rigid body over the flow strengths
tested. For this structure, the deformation measures were constant in
time for shear rates $\dot \gamma$ up to $0.15$ and TG flow rate $V_0$
up to $0.25$ (Figure~\ref{fig:BC3_shear} and Figure~\ref{fig:BC3_TG}).
In contrast, a $25\%$--$50 \%$ smaller value for $\dot \gamma$ or $V_0$ was
required to observe nonrigid deformations for the BC (i) and BC (ii)
cases.

@@

The characteristics of the self-assembly of amphiphilic JP into
onion-like dendrimersomes like for BC (ii) were previously studied by molecular dynamics
simulations~\cite{C9NR05885K}. The molecular dynamics simulations use an
anisotropic pair potential to describe the particle interactions
where a harmonic form describes the repulsive part and an anisotropic
attractive part describes the hydrophobic interactions~\cite{HongCacciutoLuijtenGranick2008}.
We also use a pair potential $P$ in the free energy \eqref{eq:free_energy} for repulsion,
whereas the long-range, hydrophobic attraction \eqref{eq:FandTdef}
is nonadditive and comes from domain-dependent variation of
the water order parameter $u(\xx)$ in the bulk.

@@

The results reported here provide an inroad modeling framework for hydrodynamics of active colloids
\cite{Meredithetal2022,McGlassonBradley2021,Vutukuri2020,Mallory2017}. 
The present study also helps us understand the rheology of JP oligomers that may be realized in the experiments.
We are now extending this study to three dimensional systems with more realistic features such as size distributions of JPs
and thermal fluctuations\cite{kohl-cor-che-vee22}.
%
 %it has been included in our future goals\cite{kohl-cor-che-vee22}.
From a numerical perspective,
it is straightforward to include
random perturbations in the particle shape and
boundary condition that mimick interfacial properties
found in lab conditions\cite{Bradley2016,Bradley2017,Zarzaretal2015,doi:10.1021/la503455h}.
